---
title: "James"
author: "student"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 2:

Data analysis. A criminologist studied the relationship between level of education and crime rate. He collected data from 84 medium-sized US counties. Two variables were measured: X- the percentage of individuals having at least a high-school diploma; and Y - the crime rate (crimes reported per 100,000 residents) in the previous year. The complete data is in "crimerate.txt".

```{r}
# Load the dataset
data <- read.table("crimerate.txt", header = TRUE)

# Check the structure of the dataset
str(data)

# Summary statistics
summary(data)

x <- data$pct.diploma
y <- data$rate
```

## (a) Calculate the least squares estimators: hat beta\_{0} , hat beta\_{1} Write down the fitted regression line, and add the line to the scatterplot of the data. Interpret hat beta\_{0} and hat beta\_{1} .

```{r}
model <- lm(y~x, data=data)
beta0 <- model$coefficients[1]
beta1 <- model$coefficients[2]
cat("the least squares estimator : hat beta_{0} is:",beta0,"\n")
cat("the least squares estimator : hat beta_{1} is:",beta1,"\n")


# Fitted regression line
fitted_line <- paste("rate = ", round(beta0, 2), " + ", round(beta1, 2), " * pct.diploma")
cat("The fitted regression line is:",fitted_line)
plot(x,y,xlab = "the percentage of individuals having at least a high-school diploma",
     ylab = "crime rate (crimes reported per 100,000 residents) in the previous year")
abline(lm(y~x),col="blue")
```

## Interpretation:

$beta_{0}$ (Intercept): The estimated intercept of the regression line is approximately 20,517.6. This value represents the predicted crime rate (rate) when the percentage of individuals having at least a high school diploma (pct.diploma) is zero. However, this interpretation may not be meaningful in the context of the data because it's unlikely that a county would have zero high school graduates. In practical terms, the intercept may not have a direct interpretation but is essential for determining the starting point of the regression line.

$beta_{1}$ (Slope): The estimated slope of the regression line is approximately -170.5752. This value represents the change in the predicted crime rate for a one-unit change in the percentage of high school graduates. Specifically, for each additional percentage point increase in the percentage of high school graduates, the predicted crime rate decreases by approximately 170.5752 crimes per 100,000 residents. Conversely, for each percentage point decrease in the percentage of high school graduates, the predicted crime rate increases by the same amount.

## (b) Calculate the estimated standard errors for the LS estimators hat beta\_{0} hat beta\_{1} respectively.

```{r}
# Standard errors of coefficients
se_beta_0 <- summary(model)$coefficients[1, 2]
se_beta_1 <- summary(model)$coefficients[2, 2]

se_beta_0
se_beta_1

```

# Assume the classical Normal error model for the rest of the problem.

## (c) Test whether or not there is a linear association between crime rate and per- centage of high school graduates at significance level 0.05. State the null and alternative hypotheses, the test statistic, its null distribution, the decision rule and the conclusion.

```{r}
# Hypothesis test for linear association
summary(model)

# Critical t value (df =82 , p)
df <- nrow(data) -1 
p <- 0.05/2
cat("The critical t-value at degree of freedom of",df,"is:",qt(p,df))
```

### State the null and alternative hypotheses:

Null Hypothesis ($H_0$): There is no linear association between crime rate and the percentage of high school graduates. Mathematically, this is represented as $H_0: \beta_1 = 0$, where $\beta_1$ is the slope coefficient.

Alternative Hypothesis ($H_1$): There is a linear association between crime rate and the percentage of high school graduates. Mathematically, this is represented as $H_1: \beta_1 \neq 0$.



### Test Statistic (t-value):

The t-value for the slope coefficient ($\beta_1$) is approximately -4.103.

Significance Level ($\alpha$):  0.05.


### Decision Rule: 

At the 0.05 significance level, I would compare the absolute value of the calculated t-statistic (-4.103) to the critical value from the t-distribution with 82 degrees of freedom (the residual degrees of freedom).
If the absolute t-statistic is greater than the critical value, I would reject the null hypothesis.
If the absolute t-statistic is not greater than the critical value, I would fail to reject the null hypothesis.

### Conclusion:
Based on the results, the absolute t-statistic (4.103) is greater than the critical value (-1.98896) from the t-distribution at a 0.05 significance level. Therefore, I would reject the null hypothesis.

### Conclusion: 
There is sufficient evidence to conclude that there is a significant linear association between crime rate and the percentage of high school graduates at the 0.05 significance level.

In simpler terms, it appears that the percentage of high school graduates has a statistically significant impact on the crime rate in the studied counties.



## (d) Construct a 95% confidence interval for beta\_{1} . Interpret your confidence interval.
```{r}
# Confidence interval for beta_1
conf_int_beta_1 <- confint(model, level = 0.95)
conf_int_beta_1

```

### Intercept (95% Confidence Interval):

For the intercept ($\beta_0$): We are 95% confident that the true intercept value falls within the interval [13997.32, 27037.88]. This interval represents the possible values of the intercept that are consistent with our data and model.

For the slope ($\beta_1$): We are 95% confident that the true slope coefficient value falls within the interval [-253.28, -87.87]. This interval represents the possible values of the slope that are consistent with our data and model.


## (e) Construct a 99% confidence interval for the mean crime rate for counties with percentage of high school graduates being 80. Interpret your confidence interval.
```{r}
# Create a new dataset with pct.diploma = 80 for prediction
new_data <- data.frame(pct.diploma = rep(80, nrow(data))) # Create a dataset with 84 rows, all with pct.diploma = 80

# Predict mean crime rates for the new dataset
pred <- predict(model, newdata = new_data, interval = "confidence", level = 0.99)
pred


```

### Interpretation:

The confidence intervals provide a range of values within which we are highly confident that the true mean crime rate lies.
For example, for the first row, the predicted mean crime rate for counties with pct.diploma = 80 is approximately 7895.036 per 100,000 residents. 
The 99% confidence interval for this prediction ranges from 7050.356 to 8739.716.

